---
title: "p8105_hw2_eb3516"
author: "Liz Bekerman"
output: github_document
date: "2023-10-02"
---

Setup code to load libraries & open dataset.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r load_libraries}
library(tidyverse)
library(readxl)
```

#Problem 1

We clean the 538 `pols` data, which provides information on the number of national politicians who are democratic or republican at any given time. There are some values for which `prez_gop` is `2` -- these are months in which Ford became president following Nixon's resignation. In the new `president` variable created as part of our data cleaning, we code these as `gop` (same values as when `prez_gop` is `1`).

```{r clean_538_pols}
month_df =
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )

pols = 
  read_csv("./fivethirtyeight_datasets/pols-month.csv") |>
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -starts_with("prez")) 
```

We also clean the 538 `snp` data, which contains information related to Standard & Poor's stock market index.

```{r clean_538_snp}
snp = 
  read_csv(
    "./fivethirtyeight_datasets/snp.csv",
    col_types = cols(date = col_date(format = "%m/%d/%y"))) |>
  separate(date, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    year = if_else(year > 2023, year - 100, year)) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, close) 
```

Finally, we tidy the `unemployment` data so that it can be merged with the `pols` and `snp` datasets.

```{r clean_538_unemp}
unemployment = 
  read_csv("./fivethirtyeight_datasets/unemployment.csv") |>
  rename(year = Year) |>
  pivot_longer(
    Jan:Dec, 
    names_to = "month_abb",
    values_to = "unemployment"
  ) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, unemployment)
```


























